---
title: "BIST8130 - Final Proejct Codings"
date: "11/22/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,dpi=300)
```

```{r library,warning=FALSE,message=FALSE}
library(tidyverse)
library(corrplot)
library(leaps)
library(performance)
library(MASS)
```

### Step 1: Data Preprocessing

After importing the csv file containing the County Demographic Information (CDI) data, we notice that crimes, physicians, and hospital beds are given as numbers, while other info are given as proportions. We therefore compute the number of crimes, physicians, and hospital beds per 1000 people.

```{r preprocess}
cdi_data = read_csv("./data/cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    density = as.numeric(pop)/as.numeric(area),
    crime_rate_1000 = 1000 * crimes/pop) %>% # Compute number of crimes per 1000 people. 
  dplyr::select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)

knitr::kable(head(cdi_data))
```


### Step 2 - Exploratory Analysis

We then take a closer look of each variables, calculate the pairwise correlations between variables, and list all the correlations between the crime rate (our interest) and all other variables.

```{r exploration}
cdi_data_exp = cdi_data %>%
  dplyr::select(-id,-cty,-state, -cty_state) 

cdi_data_original = cdi_data_exp

cdi_data_exp = cdi_data_exp %>%
  mutate(
    area = ifelse(area %in% boxplot.stats(cdi_data_exp$area)$out,NA,area),
    pop = ifelse(pop %in% boxplot.stats(cdi_data_exp$pop)$out,NA,pop),
    crime_rate_1000 = ifelse(crime_rate_1000 %in% boxplot.stats(cdi_data_exp$crime_rate_1000)$out,NA,crime_rate_1000),
    poverty = ifelse(poverty %in% boxplot.stats(cdi_data_exp$poverty)$out,NA,poverty)
  ) %>%
  na.omit()


par(mfrow=c(2,3))
boxplot(cdi_data_exp$area,main="Area")
boxplot(cdi_data_exp$pop,main="Population")
boxplot(cdi_data_exp$pop18,main="Population 18-34")
boxplot(cdi_data_exp$pop65,main="Population 65+")
boxplot(cdi_data_exp$hsgrad,main="Highschool grads")
boxplot(cdi_data_exp$bagrad,main="Bachelor's grads")

par(mfrow=c(2,3))
boxplot(cdi_data_exp$poverty,main="Poverty Rate")
boxplot(cdi_data_exp$unemp,main="Unemployment Rate")
boxplot(cdi_data_exp$pcincome,main="Income Per Capita")
boxplot(cdi_data_exp$totalinc,main="Income Total")
boxplot(cdi_data_exp$docs_rate_1000,main="Active Physicians")
boxplot(cdi_data_exp$beds_rate_1000,main="Hospital Beds")

par(mfrow=c(1,1))

ggplot(cdi_data,aes(region)) + 
  geom_histogram(binwidth = 0.5) +
  theme_classic() +
  xlab("Region")+
  ylab("Count")

boxplot(cdi_data_exp$crime_rate_1000,main="Crime Rate",horizontal = TRUE)

# data exploratory
pairs(cdi_data_exp)

# correlation plot
cdi_data_cor = cor(cdi_data_exp)
corrplot(cdi_data_cor, type = "upper", diag = FALSE)
 
crime_1000_cor = data.frame(cdi_data_cor) %>% 
  dplyr::select("Crime Rate (Per 1000)" = crime_rate_1000) %>% 
  t()

knitr::kable(crime_1000_cor,digits = 2) 

```


### Remove outliers and high leverage point

```{r}

cdi_data_clean = cdi_data[cdi_data$area >= quantile(cdi_data$area,0.005) & cdi_data$area <= quantile(cdi_data$area,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop >= quantile(cdi_data_clean$pop,0.005) & cdi_data_clean$pop <= quantile(cdi_data_clean$pop,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop18 >= quantile(cdi_data_clean$pop18,0.005) & cdi_data_clean$pop18 <= quantile(cdi_data_clean$pop18,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop65 >= quantile(cdi_data_clean$pop65,0.005) & cdi_data_clean$pop65 <= quantile(cdi_data_clean$pop65,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$hsgrad >= quantile(cdi_data_clean$hsgrad,0.005) & cdi_data_clean$hsgrad <= quantile(cdi_data_clean$hsgrad,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$bagrad >= quantile(cdi_data_clean$bagrad,0.005) & cdi_data_clean$bagrad <= quantile(cdi_data_clean$bagrad,0.995),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$poverty >= quantile(cdi_data_clean$poverty,0.005) & cdi_data_clean$poverty <= quantile(cdi_data_clean$poverty,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$unemp >= quantile(cdi_data_clean$unemp,0.005) & cdi_data_clean$unemp <= quantile(cdi_data_clean$unemp,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pcincome >= quantile(cdi_data_clean$pcincome,0.005) & cdi_data_clean$pcincome <= quantile(cdi_data_clean$pcincome,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$totalinc >= quantile(cdi_data_clean$totalinc,0.005) & cdi_data_clean$totalinc <= quantile(cdi_data_clean$totalinc,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$docs_rate_1000 >= quantile(cdi_data_clean$docs_rate_1000,0.005) & cdi_data_clean$docs_rate_1000 <= quantile(cdi_data_clean$docs_rate_1000,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.005) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.005) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.995),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$density >= quantile(cdi_data_clean$density,0.005) & cdi_data_clean$density <= quantile(cdi_data_clean$density,0.995),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$crime_rate_1000 >= quantile(cdi_data_clean$crime_rate_1000,0.005) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$crime_rate_1000,0.995),]


```

```{r}

par(mfrow=c(3,4))
boxplot(cdi_data_clean$area,main="Area")
boxplot(cdi_data_clean$pop,main="Population")
boxplot(cdi_data_clean$pop18,main="Population 18-34")
boxplot(cdi_data_clean$pop65,main="Population 65+")
boxplot(cdi_data_clean$hsgrad,main="Highschool grads")
boxplot(cdi_data_clean$bagrad,main="Bachelor's grads")

boxplot(cdi_data_clean$poverty,main="Poverty Rate")
boxplot(cdi_data_clean$unemp,main="Unemployment Rate")
boxplot(cdi_data_clean$pcincome,main="Income Per Capita")
boxplot(cdi_data_clean$totalinc,main="Income Total")
boxplot(cdi_data_clean$docs_rate_1000,main="Active Physicians")
boxplot(cdi_data_clean$beds_rate_1000,main="Hospital Beds")

```



### Model construction

Data used for building model:

```{r}
cdi_model = cdi_data %>% 
  dplyr::select(-id,-cty_state, -cty,-state) %>% 
  mutate(region = factor(region))

```


```{r}


full.fit = lm(crime_rate_1000 ~ ., data = cdi_model)
summary(full.fit) %>% 
  broom::tidy() %>%
  mutate(p_rank = rank(p.value))


backward = step(full.fit, direction='backward') %>%  broom::tidy() %>%  rename(backward = "term")


both = step(full.fit, direction = "both") %>% broom::tidy() %>% rename(stepwise = "term")

bind_cols(backward[-1,1],both[-1,1]) %>% knitr::kable()

```

## Criteria based selection

pop pop18 hsgrad bagrad poverty pcincome totalinc region beds_rate_1000 density

```{r}
sb = regsubsets(crime_rate_1000 ~ ., data = cdi_model, nvmax = 14)
sumsb = summary(sb) # pop pop18 hsgrad bagrad poverty pcincome totalinc region beds_rate_1000 density
sumsb

# plot of Cp and Adj-R2 as functions of parameters
par(mfrow=c(1,2))
plot(2:15, sumsb$cp, xlab="No. of parameters", ylab="Cp Statistic") 
abline(0,1)

plot(2:15, sumsb$adjr2, xlab="No of parameters", ylab="Adj R2")

#112
```



## Interaction

Does the relationship between the crime_rate_1000 and poverty vary by region status?

```{r}
ggplot(cdi_data, aes(x = poverty, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact = lm(crime_rate_1000 ~ poverty*region, data = cdi_data)
summary(interact)
```

## test

```{r}
fit_test = lm(crime_rate_1000 ~  
                  pop + pop18 +hsgrad + bagrad+ pcincome + totalinc + region  +
                  beds_rate_1000 + density + poverty +
                  poverty*pcincome, data = cdi_model)
summary(fit_test)


check_collinearity(fit_test)
```

pcincome*bagrad significant
poverty*income significnat, hsgrad, bagrad, poverty insignificant

## find outlier

```{r}
plot(fit_test, which = 4) 

cdiOut = cdi_model[-c(1,6),]

# fit model with and without influential points
with = fit_test 

without = lm(crime_rate_1000 ~  
                  pop + pop18   
                  + pcincome + totalinc + region +
                  beds_rate_1000 + density + 
                  poverty*pcincome, data = cdiOut)

summary(with); summary(without) 

par(mfrow = c(2,2))
plot(without)

boxcox(without)
```


Method:

1. Data processing: 

* Summarize `docs`,`beds`and `crimes` rate per 1000 population

* Create a new variable, `density`, which is population divided by area

2. Exploratory analysis:

* Calculate the pairwise correlations between variables

* List all the correlations between the crime rate (our interest) and all other variables

3. Remove outliers and high leverage points

Use percentile to detect potential outliers and high leverage points. Due to the dataset size, we remove rows containing the smallest and largest 0.2% for each variables.

4. Training/Testing set split:
Randomly split the dataset into training and testing sets. 90% is training set while 10% is testing. This step aims to support model assessment and avoid overestimation.

5. Model construction:

* Select variables using stepwise regression and criteria based procedure

* Build model using the variables we selected 

* Plot interaction effects and add interaction terms

* Diagnose and transform the models

6. Cross validation

Cross validate on each model and get the model with the lowest RMSE.

7. Model assessment


Reference:
As Numerous new questions emerging during our discussion, our group explored materials below to solve them.

1. Question: Do we still need a test set when using k-fold cross-validation?
Source:
https://stats.stackexchange.com/questions/225949/do-we-need-a-test-set-when-using-k-fold-cross-validation
https://datascience.stackexchange.com/questions/80310/is-a-test-set-necessary-after-cross-validation-on-training-set


2. Question: How to achieve build test set & predict
Source:
https://www.ritchieng.com/machine-learning-evaluate-linear-regression-model/
https://campus.datacamp.com/courses/machine-learning-with-caret-in-r/regression-models-fitting-them-and-evaluating-their-performance?ex=8

3.Question: How to evaluate continuous by continuous interactions
Source: Continuous by Continuous Interactions, Joel S Steele http://web.pdx.edu/~joel8/resources/ConceptualPresentationResources/ContinuousByContinousInteractions_walkthrough_v2.pdf


Other reference:
* Friedman, Jerome, Trevor Hastie, and Robert Tibshirani. The elements of statistical learning. Vol. 1. No. 10. New York: Springer series in statistics, 2001.
