---
title: "BIST8130 - Final Proejct Codings"
date: "11/22/2021"
fontsize: 11pt
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,dpi=300)
```

```{r library,warning=FALSE,message=FALSE}
library(tidyverse)
library(corrplot)
library(leaps)
library(performance)
library(MASS)
library(caret)

```

## Step 1: Data Preprocessing

After importing the csv file containing the County Demographic Information (CDI) data, we notice that crimes, physicians, and hospital beds are given as numbers, while other info are given as proportions. We therefore compute the number of crimes, physicians, and hospital beds per 1000 people.

```{r preprocess}
cdi_data = read_csv("./data/cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    density = as.numeric(pop)/as.numeric(area),
    crime_rate_1000 = 1000 * crimes/pop) %>% # Compute number of crimes per 1000 people. 
  dplyr::select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)

knitr::kable(head(cdi_data))
```


## Step 2 - Exploratory Analysis

We then take a closer look of each variables, calculate the pairwise correlations between variables, and list all the correlations between the crime rate (our interest) and all other variables.

```{r exploration}
cdi_data_exp = cdi_data %>%
  dplyr::select(-id,-cty,-state, -cty_state) 

cdi_data_original = cdi_data_exp

cdi_data_exp = cdi_data_exp %>%
  mutate(
    area = ifelse(area %in% boxplot.stats(cdi_data_exp$area)$out,NA,area),
    pop = ifelse(pop %in% boxplot.stats(cdi_data_exp$pop)$out,NA,pop),
    crime_rate_1000 = ifelse(crime_rate_1000 %in% boxplot.stats(cdi_data_exp$crime_rate_1000)$out,NA,crime_rate_1000),
    poverty = ifelse(poverty %in% boxplot.stats(cdi_data_exp$poverty)$out,NA,poverty)
  ) %>%
  na.omit()


par(mfrow=c(2,3))
boxplot(cdi_data_exp$area,main="Area")
boxplot(cdi_data_exp$pop,main="Population")
boxplot(cdi_data_exp$pop18,main="Population 18-34")
boxplot(cdi_data_exp$pop65,main="Population 65+")
boxplot(cdi_data_exp$hsgrad,main="Highschool grads")
boxplot(cdi_data_exp$bagrad,main="Bachelor's grads")

par(mfrow=c(2,3))
boxplot(cdi_data_exp$poverty,main="Poverty Rate")
boxplot(cdi_data_exp$unemp,main="Unemployment Rate")
boxplot(cdi_data_exp$pcincome,main="Income Per Capita")
boxplot(cdi_data_exp$totalinc,main="Income Total")
boxplot(cdi_data_exp$docs_rate_1000,main="Active Physicians")
boxplot(cdi_data_exp$beds_rate_1000,main="Hospital Beds")

par(mfrow=c(1,1))

ggplot(cdi_data,aes(region)) + 
  geom_histogram(binwidth = 0.5) +
  theme_classic() +
  xlab("Region")+
  ylab("Count")

boxplot(cdi_data_exp$crime_rate_1000,main="Crime Rate",horizontal = TRUE)

# data exploratory
pairs(cdi_data_exp)

# correlation plot
cdi_data_cor = cor(cdi_data_exp)
corrplot(cdi_data_cor, type = "upper", diag = FALSE)
 
crime_1000_cor = data.frame(cdi_data_cor) %>% 
  dplyr::select("Crime Rate (Per 1000)" = crime_rate_1000) %>% 
  t()

knitr::kable(crime_1000_cor,digits = 2) 

```


## Remove outliers and high leverage point

```{r}
# Remove high leverage points

cdi_data_clean = cdi_data[cdi_data$area >= quantile(cdi_data$area,0.002) & cdi_data$area <= quantile(cdi_data$area,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop >= quantile(cdi_data_clean$pop,0.002) & cdi_data_clean$pop <= quantile(cdi_data_clean$pop,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop18 >= quantile(cdi_data_clean$pop18,0.002) & cdi_data_clean$pop18 <= quantile(cdi_data_clean$pop18,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop65 >= quantile(cdi_data_clean$pop65,0.002) & cdi_data_clean$pop65 <= quantile(cdi_data_clean$pop65,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$hsgrad >= quantile(cdi_data_clean$hsgrad,0.002) & cdi_data_clean$hsgrad <= quantile(cdi_data_clean$hsgrad,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$bagrad >= quantile(cdi_data_clean$bagrad,0.002) & cdi_data_clean$bagrad <= quantile(cdi_data_clean$bagrad,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$poverty >= quantile(cdi_data_clean$poverty,0.002) & cdi_data_clean$poverty <= quantile(cdi_data_clean$poverty,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$unemp >= quantile(cdi_data_clean$unemp,0.002) & cdi_data_clean$unemp <= quantile(cdi_data_clean$unemp,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pcincome >= quantile(cdi_data_clean$pcincome,0.002) & cdi_data_clean$pcincome <= quantile(cdi_data_clean$pcincome,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$totalinc >= quantile(cdi_data_clean$totalinc,0.002) & cdi_data_clean$totalinc <= quantile(cdi_data_clean$totalinc,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$docs_rate_1000 >= quantile(cdi_data_clean$docs_rate_1000,0.002) & cdi_data_clean$docs_rate_1000 <= quantile(cdi_data_clean$docs_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$density >= quantile(cdi_data_clean$density,0.002) & cdi_data_clean$density <= quantile(cdi_data_clean$density,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$crime_rate_1000 >= quantile(cdi_data_clean$crime_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$crime_rate_1000,0.998),]


```


```{r}

par(mfrow=c(3,4))
boxplot(cdi_data_clean$area,main="Area")
boxplot(cdi_data_clean$pop,main="Population")
boxplot(cdi_data_clean$pop18,main="Population 18-34")
boxplot(cdi_data_clean$pop65,main="Population 65+")
boxplot(cdi_data_clean$hsgrad,main="Highschool grads")
boxplot(cdi_data_clean$bagrad,main="Bachelor's grads")

boxplot(cdi_data_clean$poverty,main="Poverty Rate")
boxplot(cdi_data_clean$unemp,main="Unemployment Rate")
boxplot(cdi_data_clean$pcincome,main="Income Per Capita")
boxplot(cdi_data_clean$totalinc,main="Income Total")
boxplot(cdi_data_clean$docs_rate_1000,main="Active Physicians")
boxplot(cdi_data_clean$beds_rate_1000,main="Hospital Beds")

```


## Training/Test set split

```{r}
cdi_data_clean = cdi_data_clean %>% 
  dplyr::select(-id,-cty_state, -cty,-state) %>% 
  mutate(region = factor(region))

set.seed(1)
dt = sort(sample(nrow(cdi_data_clean), nrow(cdi_data_clean)*.9))
train_data = cdi_data_clean[dt,]
test_data = cdi_data_clean[-dt,]

```

## Model construction

Data used for building model:

```{r}

cdi_model = train_data

```


### Stepwise regression

```{r}

full.fit = lm(crime_rate_1000 ~ ., data = cdi_model)
summary(full.fit) %>% 
  broom::tidy() %>%
  mutate(p_rank = rank(p.value))


backward = step(full.fit, direction='backward') %>%  broom::tidy() %>%  rename(backward = "term")


both = step(full.fit, direction = "both") %>% broom::tidy() %>% rename(stepwise = "term")

bind_cols(backward[-1,1],both[-1,1]) %>% knitr::kable()

```

### Criteria based selection


```{r}
sb = regsubsets(crime_rate_1000 ~ ., data = cdi_model, nvmax = 14)
sumsb = summary(sb) # pop pop18 hsgrad bagrad poverty pcincome totalinc region beds_rate_1000 density
sumsb

# plot of Cp and Adj-R2 as functions of parameters
par(mfrow=c(1,2))
plot(2:15, sumsb$cp, xlab="No. of parameters", ylab="Cp Statistic") 
abline(0,1)

plot(2:15, sumsb$adjr2, xlab="No of parameters", ylab="Adj R2")
```

According to the output, we determine that the number of variables should be above 12 because $C_p \leq p$. Based on this analysis, we find that `unemp` could also be selected.

### Discussion

We need to remove totalinc, because it can be replaced. totalinc = pcincome * pop.

## Model building from the vairables we selected

```{r}
fit_nest = lm(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000+ docs_rate_1000 + density, data = cdi_model)
summary(fit_nest)

# remove docs
fit_nest = lm(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density, data = cdi_model)
summary(fit_nest)


```

Our first model: crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density
                  
## Add Interaction term: poverty+income

According to Census Bureau, the number of persons below the official government poverty level was 33.6 million in 1990, representing 13.5 percent of the Nation's population. Thus, we can use this criteria to divide `poverty` into two category: higher than national poverty rate and lower than national poverty rate.

```{r}
poverty_status = cdi_model %>% 
  mutate(national_poverty = if_else(poverty > 13.5, "higher", "lower"))

ggplot(poverty_status, aes(x = pcincome, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) +
  labs(
    title = "Crime Rate and Per Capita Income by Poverty Status",
    x = "Per Capita Income",
    y = "Crime Rate "
  )
```


```{r}
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1)

check_collinearity(fit_int1)

# remove density
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1)

check_collinearity(fit_int1)

# remove bagrad
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1)

check_collinearity(fit_int1)
```


### diagnose
```{r}
plot(fit_int1, which = 1)
plot(fit_int1, which = 2)
boxcox(fit_int1)
```

box close to 0.5~1. Try y^0.5 transformation

### transformation

```{r}

cdi_model_trans = cdi_model %>% 
  mutate(
    y_sqrt = sqrt(crime_rate_1000)
  )

fit_int1_trans = lm(y_sqrt ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model_trans)
summary(fit_int1_trans)

# diagnose
plot(fit_int1_trans, which = 1) #worse
plot(fit_int1_trans, which = 2) 



```

Transformed is worse.

The second model: crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome

## Add interaction term: pcincome + bagrad

According to Census Bureau, national percent of persons 25 years old or older with bachelor’s degrees is 20.8%. Thus, we can use this criteria to divide `bagrad` into two category: higher than national `bagrad` and lower than national `bargrad`.

```{r}
bagrad_status = cdi_model %>% 
  mutate(national_bagrad = if_else(bagrad > 20.8, "higher", "lower"))

ggplot(bagrad_status, aes(x = pcincome, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) +
  ylim(0,150) +
  labs(
    title = "Crime Rate and Per Capita Income by Percent Bachelor’s Degrees Status",
    x = "Per Capita Income",
    y = "Crime Rate"
  )

```

```{r}
fit_int2 = lm(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad, data = cdi_model)
summary(fit_int2)

check_collinearity(fit_int2)
```

### diagnose

```{r}

plot(fit_int2, which = 1)
plot(fit_int2, which = 2)
boxcox(fit_int2)
```


box close to 0.5~1. Try y^0.5 transformation


### transformation
```{r}

fit_int2_trans = lm(y_sqrt ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad, data = cdi_model_trans)
summary(fit_int2_trans)

# diagnose
par(mfrow=c(2,2))
plot(fit_int2_trans, which = 1) # worse
plot(fit_int2_trans, which = 2) 



```

Transformed is worse.

The third model: crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad
     
                  
## Cross validation

### model 1

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train1 = train(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density,data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)
print(model_train1)

```


### model 2

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train2 = train(crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

summary(model_train2)


```

### model 3

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train3 = train(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad,  data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)
summary(model_train3)

```

### Compare RMSE

```{r}

model <- c("1", "2", "3")

RMSE <- c(round(model_train1$results$RMSE, 2),  round(model_train2$results$RMSE,2),
          round(model_train3$results$RMSE, 2))

R_sq <- c(round(model_train1$results$Rsquared, 3),
          round(model_train2$results$Rsquared, 3),
          round(model_train3$results$Rsquared, 3))

RMSE_table <- data.frame(model, RMSE, R_sq)

coefs_1 = model_train1$finalModel$coefficients
names_1 = model_train1$finalModel$xNames

knitr::kable(RMSE_table)
```

The second model has the lowest RMSE.

### Model Assessment on testing set

```{r}

test_data = test_data %>%
  mutate(
    y = crime_rate_1000,
    y_model_1 = predict(model_train1,test_data),
    y_model_2 = predict(model_train2,test_data),
    y_model_3 = predict(model_train3,test_data))

RMSPE_1 = sqrt(mean((test_data$y-test_data$y_model_1)^2))
RMSPE_2 = sqrt(mean((test_data$y-test_data$y_model_2)^2))
RMSPE_3 = sqrt(mean((test_data$y-test_data$y_model_3)^2))



model_assessment = 
  tibble(
    RMSPE_1 = round(RMSPE_1,2),
    RMSPE_2 = round(RMSPE_2,2),
    RMSPE_3 = round(RMSPE_3,2)) %>% 
  pivot_longer(RMSPE_1:RMSPE_3,
               names_to = "model", 
               names_prefix = "RMSPE_",
               values_to = "RMSPE") %>%
  left_join(RMSE_table,by="model") %>%
  dplyr::select(Model=model,R_square = R_sq,RMSE,RMSPE)

knitr::kable(model_assessment)
```