---
title: "BIST8130 - Final Project Report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

In this project, we aim to build regression models based on a set of demographic variables to estimate county-level crime rates. After an exploratory analysis of the variables on their distributions and correlations, we derived more meaningful variables by manipulating the existing ones, removed outlier values, and implement several variable selection methods. Using the selected variables, we trained a linear regression model, elaborate on the model by adding several interactive terms, and did cross-validations. All 3 resulting models have achieved a good estimation of the training set. The first model has the best prediction on new data (in the testing set), whereas the other two models, despite a better performance of the training data, may have the problem of overfitting.

## Team Member

Mengfan Luo (ml4701), Yushan Wang (yw3772), Jing Lyu(jl6049), Yiqun Jin (yj2686), Mingkuan Xu (mx2262).
 
## Introduction

Over the last three decades, crime has become a major public concern in the US arousing massive political discussion and public expenditure[1]. Crime rates in major cities experienced a general rise from the 1960s to 1990s, with two peaks observed in 1980 and in early 1990s[2]. Despite extensive attention across the nation, factors influencing crime trends were not yet made clear[1]. In this project, we examined crime rates and potential affecting factors in 440 US counties from "County Demographic Information" (CDI) data set, and construct a multiple linear regression model to predict the crime rates.




## Methods

### Dataset discription

We analyzed data from the “County Demographic Information” (CDI) data set, which contains characteristics of 440 counties in the United States collected from 1990-1992. The primary goal of this investigation is to develop insight relevant to predicting the crime rate in counties. 

### Model building method

1. Data preprocessing: 

Considering transformation of vairables in order to extract interpretable information form correlated predictors.

2. Exploratory analysis:

* Calculate the pairwise correlations between variables

* List all the correlations between the crime rate (our interest) and all other variables

3. Training/Testing set split:
Randomly split the dataset into training and testing sets. 90% is training set while 10% is testing. This step aims to support model assessment and avoid overestimation.

4. Remove outliers and high leverage points

Use percentile to detect potential outliers and high leverage points. Due to the dataset size, we remove rows containing the smallest and largest 0.2% for each variables.



5. Model construction:

* Select variables using stepwise regression and criteria based procedure

* Build model using the variables we selected 

* Plot interaction effects and add interaction terms

* Diagnose and transform the models

6. Cross validation

Cross validate on each model and get the model with the lowest RMSE.

7. Model assessment
In this section, we intend to assess the models we built and choose a final model used for future prediction. We compared the models based on three criteria: the R-square values, the root mean square error (RMSE), and the root mean square prediction error (RMSPE).  

- R square represents the proportion of the variance that can be explained by the regression model.
- RMSE measures the differences between the actual values and the predicted values in the training dataset.
- RMSPE estimates the prediction errors on new data outside the training dataset.


## Results

### Descriptive analysis

After importing the csv file containing the County Demographic Information (CDI) data, we noticed that crimes, physicians, and hospital beds are given as numbers, while other info are given as proportions. We therefore compute the number of crimes, physicians, and hospital beds per 1000 people. When considering crime rate, population density could be a key factor. Thus, we also create a new variable, `density`, which is population divided by area. 


After drawing boxplots to show the distribution of the variables, we identified several extreme values in each of the variables. These values can be treated as potential outliers to be removed in further analysis. For example, the distribution of crime rate:

```{r, fig.cap="\\label{fig:figs}boxplot of dependent variable: crime rate"}
boxplot(cdi_data_exp$crime_rate_1000,main="Crime Rate",horizontal = TRUE)
```


We then take a closer look of each variables, calculate the pairwise correlations between variables, and list all the correlations between the crime rate (our interest) and all other variables. We used correlation analysis to prove that the derived variable density is a more meaningful variable compared to area and population, given that it has a stronger association with the crime rate.

```{r preprocess, messgae = FALSE}
cdi_data = read_csv("./data/cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, 
    # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    density = as.numeric(pop)/as.numeric(area),
    crime_rate_1000 = 1000 * crimes/pop) %>% 
  # Compute number of crimes per 1000 people. 
  dplyr::select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)


cdi_data_exp = cdi_data %>%
  dplyr::select(-id,-cty,-state, -cty_state) 



```

```{r, fig.cap="\\label{fig:figs}Correlation heatmap"}
# correlation plot
cdi_data_cor = cor(cdi_data_exp)
corrplot(cdi_data_cor, type = "upper", diag = FALSE)
```
After preliminary analysis of the data, we identify several variables that might be relevant to the crime rate as listed 

```{r, echo = FALSE}
Variable <- c("area","density",	"pop", "pop18",	"pop65",	"docs_rate_1000",	"beds_rate_1000",	"crime_rate_1000",	"hsgrad",	"bagrad",	"poverty",	"unemp",	"pcincome",	"totalinc",	"region")

Meaning <- c("Land area","Population Density" ,"Esimate 1990 population", "Percent of population aged 18-34", "Percent of population aged 65+", "Number of active physicians per 1000 people", "Number of hospital beds per 1000 people", "Number of serious crimes per 1000 people", "Percent high school graduates", "Percent bachelor’s degrees", "Percent below poverty level", "Percent unemployment", "Per capita income", "Total personal income", "Geographic region")

var_info <- data.frame(Variable, Meaning)

knitr::kable(var_info, caption = "Potential Variables Relevant to the Crime Rate")
```

### Traning/Testing split

We randomly sampled 10% from the dataset and made it a testing set(n = 44). The rest is training set (n = 396).

```{r}
cdi_data = cdi_data %>% 
  dplyr::select(-id,-cty_state, -cty,-state) %>% 
  mutate(region = factor(region))

set.seed(1)
dt = sort(sample(nrow(cdi_data), nrow(cdi_data)*.9))
train_data = cdi_data[dt,]
test_data = cdi_data[-dt,]

```



### Data cleaning

Due to the dataset size, we removed rows containing the smallest and largest 0.2% for each variable. 19 data points were removed out of 396.

```{r echo=FALSE}
# Remove high leverage points

cdi_data_clean = train_data[train_data$area >= quantile(train_data$area,0.002) & train_data$area <= quantile(train_data$area,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop >= quantile(cdi_data_clean$pop,0.002) & cdi_data_clean$pop <= quantile(cdi_data_clean$pop,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop18 >= quantile(cdi_data_clean$pop18,0.002) & cdi_data_clean$pop18 <= quantile(cdi_data_clean$pop18,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop65 >= quantile(cdi_data_clean$pop65,0.002) & cdi_data_clean$pop65 <= quantile(cdi_data_clean$pop65,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$hsgrad >= quantile(cdi_data_clean$hsgrad,0.002) & cdi_data_clean$hsgrad <= quantile(cdi_data_clean$hsgrad,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$bagrad >= quantile(cdi_data_clean$bagrad,0.002) & cdi_data_clean$bagrad <= quantile(cdi_data_clean$bagrad,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$poverty >= quantile(cdi_data_clean$poverty,0.002) & cdi_data_clean$poverty <= quantile(cdi_data_clean$poverty,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$unemp >= quantile(cdi_data_clean$unemp,0.002) & cdi_data_clean$unemp <= quantile(cdi_data_clean$unemp,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pcincome >= quantile(cdi_data_clean$pcincome,0.002) & cdi_data_clean$pcincome <= quantile(cdi_data_clean$pcincome,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$totalinc >= quantile(cdi_data_clean$totalinc,0.002) & cdi_data_clean$totalinc <= quantile(cdi_data_clean$totalinc,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$docs_rate_1000 >= quantile(cdi_data_clean$docs_rate_1000,0.002) & cdi_data_clean$docs_rate_1000 <= quantile(cdi_data_clean$docs_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$density >= quantile(cdi_data_clean$density,0.002) & cdi_data_clean$density <= quantile(cdi_data_clean$density,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$crime_rate_1000 >= quantile(cdi_data_clean$crime_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$crime_rate_1000,0.998),]


```


### Model construction

* Select variables using stepwise regression and criteria based procedure

* Build model using the variables we selected 

* Plot interaction effects and add interaction terms

* Diagnose and transform the models




## Model construction

Data used for building model:

```{r}

cdi_model = cdi_data_clean

```


### Stepwise regression

```{r}

full.fit = lm(crime_rate_1000 ~ ., data = cdi_model)
summary(full.fit) %>% 
  broom::tidy() %>%
  mutate(p_rank = rank(p.value))

backward = step(full.fit, direction='backward') %>%  broom::tidy() %>%  rename(backward = "term")


both = step(full.fit, direction = "both") %>% broom::tidy() %>% rename(stepwise = "term")


```

Variables chosen from stepwise regression:

```{r}
bind_cols(backward[-1,1],both[-1,1]) %>% knitr::kable(caption = "Vairable selected from stepwise regression")

```

### Criteria based selection


```{r}
sb = regsubsets(crime_rate_1000 ~ ., data = cdi_model, nvmax = 14)
sumsb = summary(sb) # pop pop18 hsgrad bagrad poverty pcincome totalinc region beds_rate_1000 density
```

```{r}
coef(sb, id = 12)

```


```{r, fig.cap="\\label{fig:figs}Subset selection for best parameter numbers"}
par(mfrow=c(1,2))
plot(2:15, sumsb$cp, xlab="No. of parameters", ylab="Cp Statistic") 
abline(0,1)

plot(2:15, sumsb$adjr2, xlab="No of parameters", ylab="Adj R2")
```

According to the output, we determine that the number of variables should be above 12 because $C_p \leq p$. Based on this analysis, we find that `unemp` could also be selected.

### Discussion

We need to remove totalinc, because it can be replaced. totalinc = pcincome * pop.

## Model building from the vairables we selected

```{r}
fit_nest = lm(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density, data = cdi_model)
summary(fit_nest)
```

```{r, fig.cap="\\label{fig:figs}Diagnose plots of model without interaction terms"}
par(mfrow = c(2,2))
plot(fit_nest)
```


```{r,fig.cap="\\label{fig:figs}Boxcox plot of model without interaction terms" }
boxcox(fit_nest)


```

The peak of boxcox plot is close to around 0.5~1. Try $\sqrt{y}$ transformation

### transformation

```{r}
cdi_model_trans = cdi_model %>% 
  mutate(
    y_sqrt = sqrt(crime_rate_1000)
  )

fit_nest_trans = lm(y_sqrt ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density, data = cdi_model_trans)
summary(fit_nest_trans)
```


```{r, fig.cap="\\label{fig:figs}Diagnose plots of model without interaction terms"}

par(mfrow = c(2,2))
plot(fit_nest_trans)



```

Compare to the diagnose plots of untransformed model, we found that the residuals are more unevenly distributed. Therefore, transformed model is worse. We select the untransformed model.


Our first model: $$crime\_rate\_1000 = pop + pop18 + bagrad +
                  poverty + unemp \\ + pcincome + pcincome*pop + region
                  beds\_rate\_1000 + density$$
                  
## Add Interaction term: poverty+income

According to Census Bureau, the number of persons below the official government poverty level was 33.6 million in 1990, representing 13.5 percent of the Nation's population. Thus, we can use this criteria to divide `poverty` into two category: higher than national poverty rate and lower than national poverty rate.

```{r, fig.cap="\\label{fig:figs}Interaction plot of Income Per Capita and Poverty"}
poverty_status = cdi_model %>% 
  mutate(national_poverty = if_else(poverty > 13.5, "higher", "lower"))

ggplot(poverty_status, aes(x = pcincome, y = crime_rate_1000, color = national_poverty)) + 
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) +
  labs(
    title = "Crime Rate and Per Capita Income by Poverty Status",
    x = "Per Capita Income",
    y = "Crime Rate ",
    color = "Comparison with national avergae"
  )
```


```{r}
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1) %>% broom::tidy()
```



```{r}
check_collinearity(fit_int1)
```

We notice that `density`, `bagrad` are not significant

```{r}
# remove density
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1)

check_collinearity(fit_int1)

# remove bagrad
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1)

check_collinearity(fit_int1)
```


### diagnose
```{r, fig.cap="\\label{fig:figs}Diagnose plots with interaction terms:poverty*pcincome"}
par(mfrow = c(2,2))
plot(fit_int1)
```


```{r, fig.cap="\\label{fig:figs}Boxcox plot with interaction terms:poverty*pcincome"}
boxcox(fit_int1)
```

The peak of boxcox plot is close to around 0.5~1. Try $\sqrt{y}$ transformation

### transformation

```{r}
cdi_model_trans = cdi_model %>% 
  mutate(
    y_sqrt = sqrt(crime_rate_1000)
  )

fit_int1_trans = lm(y_sqrt ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model_trans)
summary(fit_int1_trans)
```


```{r, fig.cap="\\label{fig:figs}Diagnose plots with interaction terms:poverty*pcincome"}
par(mfrow = c(2,2))
plot(fit_int1_trans)
```

Compare to the diagnose plots of untransformed model, we found that the residuals are more unevenly distributed. Therefore, transformed model is worse. We select the untransformed model.

                  
Our second model: $$crime\_rate\_1000 = pop + pop18 + 
                  poverty + unemp + pcincome + \\ pcincome*pop + region +
                  beds\_rate\_1000  +
                  poverty*pcincome$$

## Add interaction term: pcincome + bagrad

According to Census Bureau, national percent of persons 25 years old or older with bachelor’s degrees is 20.8%. Thus, we can use this criteria to divide `bagrad` into two category: higher than national `bagrad` and lower than national `bargrad`.

```{r, fig.cap="\\label{fig:figs}Interaction plot of Income Per Capita and Bachelor's Degree Status"}
bagrad_status = cdi_model %>% 
  mutate(national_bagrad = if_else(bagrad > 20.8, "higher", "lower"))

ggplot(bagrad_status, aes(x = pcincome, y = crime_rate_1000, color = national_bagrad)) + 
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) +
  ylim(0,150) +
  labs(
    title = "Crime Rate and Per Capita Income by Percent Bachelor's Degrees Status",
    x = "Per Capita Income",
    y = "Crime Rate",
    color = "Comparison with national avergae"
  )

```

```{r}
fit_int2 = lm(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad, data = cdi_model)
summary(fit_int2)

check_collinearity(fit_int2)
```

### diagnose

```{r, fig.cap="\\label{fig:figs}Diagnose plots with interaction terms:pcincome*bagrad"}
par(mfrow = c(2,2))
plot(fit_int2)
```


```{r, fig.cap="\\label{fig:figs}Boxcox plot with interaction terms:pcincome*bagrad"}
boxcox(fit_int2)
```


The peak of boxcox plot is close to around 0.5~1. Try $\sqrt{y}$ transformation


### transformation
```{r}
fit_int2_trans = lm(y_sqrt ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad, data = cdi_model_trans)
summary(fit_int2_trans)
```


```{r, fig.cap="\\label{fig:figs}Diagnose plots with interaction terms:pcincome*bagrad"}
par(mfrow = c(2,2))
plot(fit_int2_trans)
``` 

Compare to the diagnose plots of untransformed model, we found that the residuals are more unevenly distributed. Therefore, transformed model is worse. We select the untransformed model.

     
Our third model: $$crime\_rate\_1000 = pop + pop18 + bagrad +
                  poverty + unemp + pcincome + \\ pcincome*pop + region +
                  beds\_rate\_1000 + density +
                  pcincome*bagrad$$                  


### 7. Cross validation

Cross validate on each model and get the model with the lowest RMSE.
## Cross validation

### model 1

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train1 = train(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density,data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)
print(model_train1)

```


### model 2

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train2 = train(crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

summary(model_train2)


```

### model 3

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train3 = train(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad,  data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)
summary(model_train3)

```

### 8. Model assessment
In this section, we intend to assess the models we built and choose a final model used for future prediction. We compared the models based on three criteria: the R-square values, the root mean square error (RMSE), and the root mean square prediction error (RMSPE).  

- R square represents the proportion of the variance that can be explained by the regression model.
- RMSE measures the differences between the actual values and the predicted values in the training dataset.
- RMSPE estimates the prediction errors on new data outside the training dataset.


* Model assessment


## Conclusion and Discussion

According to the above table, evaluated in terms of accuracy in estimating the training set, model 2 has the best performance with its leading R square and the smallest RMSE, followed by model 3 and model 1. In terms of predicting values outside the training set, model 1 has the best performance with the smallest RMSPE on the testing set. Such an inconsistency of model performance may be explained by model overfitting of the training data. Above all, although models 2 and 3 have achieved a good estimation of the training set, we will choose model 1 as the final model, given its fair RMSE and R-square values, plus excellent testing set performance.  

 

Overall, our project has several strengths. First, we did feature engineering before training the model by transforming variables using our domain knowledge to the ones more relevant to the predicted variable. For example, while area and population are not directly related to the crime rate, population density (population/area) can be more relevant. Second, we did an analysis of correlation and collinearity, which reduced the potential bias caused by confounders. Third, we did an interactive analysis and involved multiple interactive terms in our model, which to some extent represented the possible non-linear relations between the parameters and the predicted value. Finally, we separated testing set from the dataset at the very beginning, on which we evaluated the performance of several models, addressing the potential predicting errors caused by model overfitting.  

 

Meanwhile, the project also has its limitations. Essentially, we identified several wrong data points in the original dataset: for example, the population of Los Angles, an outstandingly large number, is not consistent with the number found on Wikipedia.Given that some of the data are mistaken, further works can be done on correcting the dataset using external data sources. Furthermore, our regression model considered only linear and interactive terms, while some parameters could be better fitted using polynomials or exponents. More sophisticated models can be used in the future to better estimate the data.

## Reference and Documentation

[1] Committee on Law and Justice, et al. *Understanding Crime Trends: Workshop Report*. Edited by Arthur S. Goldberger and Richard Rosenfeld, National Academies Press, 2009. Accessed 11 December 2021.

[2] Rosenfeld, R., Vogel, M. & McCuddy, T. Crime and Inflation in U. S. Cities. *J Quant Criminol* 35, 195–210 (2019). https://doi.org/10.1007/s10940-018-9377-x

[3] Friedman, Jerome, Trevor Hastie, and Robert Tibshirani. The elements of statistical learning. Vol. 1. No. 10. New York: Springer series in statistics, 2001.


As Numerous new questions emerging during our discussion, our group explored materials below to solve them.

1. Question: Do we still need a test set when using k-fold cross-validation?
Source:
https://stats.stackexchange.com/questions/225949/do-we-need-a-test-set-when-using-k-fold-cross-validation
https://datascience.stackexchange.com/questions/80310/is-a-test-set-necessary-after-cross-validation-on-training-set


2. Question: How to achieve build test set & predict
Source:
https://www.ritchieng.com/machine-learning-evaluate-linear-regression-model/
https://campus.datacamp.com/courses/machine-learning-with-caret-in-r/regression-models-fitting-them-and-evaluating-their-performance?ex=8

3.Question: How to evaluate continuous by continuous interactions
Source: Continuous by Continuous Interactions, Joel S Steele http://web.pdx.edu/~joel8/resources/ConceptualPresentationResources/ContinuousByContinousInteractions_walkthrough_v2.pdf



