---
title: "YiqunJin"
author: "Yiqun Jin"
date: "11/22/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(corrplot)
library(performance)
library(dplyr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Load Data

```{r}
cdi_data = read.csv("data/cdi.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    density = as.numeric(pop)/as.numeric(area),
    crime_rate_1000 = 1000 * crimes/pop) %>% # Compute number of crimes per 1000 people. 
  dplyr::select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)

knitr::kable(head(cdi_data))
  
```

### Remove outliers and high leverage point

```{r}
# Remove high leverage points

cdi_data_clean = cdi_data[cdi_data$area >= quantile(cdi_data$area,0.002) & cdi_data$area <= quantile(cdi_data$area,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop >= quantile(cdi_data_clean$pop,0.002) & cdi_data_clean$pop <= quantile(cdi_data_clean$pop,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop18 >= quantile(cdi_data_clean$pop18,0.002) & cdi_data_clean$pop18 <= quantile(cdi_data_clean$pop18,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop65 >= quantile(cdi_data_clean$pop65,0.002) & cdi_data_clean$pop65 <= quantile(cdi_data_clean$pop65,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$hsgrad >= quantile(cdi_data_clean$hsgrad,0.002) & cdi_data_clean$hsgrad <= quantile(cdi_data_clean$hsgrad,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$bagrad >= quantile(cdi_data_clean$bagrad,0.002) & cdi_data_clean$bagrad <= quantile(cdi_data_clean$bagrad,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$poverty >= quantile(cdi_data_clean$poverty,0.002) & cdi_data_clean$poverty <= quantile(cdi_data_clean$poverty,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$unemp >= quantile(cdi_data_clean$unemp,0.002) & cdi_data_clean$unemp <= quantile(cdi_data_clean$unemp,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pcincome >= quantile(cdi_data_clean$pcincome,0.002) & cdi_data_clean$pcincome <= quantile(cdi_data_clean$pcincome,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$totalinc >= quantile(cdi_data_clean$totalinc,0.002) & cdi_data_clean$totalinc <= quantile(cdi_data_clean$totalinc,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$docs_rate_1000 >= quantile(cdi_data_clean$docs_rate_1000,0.002) & cdi_data_clean$docs_rate_1000 <= quantile(cdi_data_clean$docs_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$density >= quantile(cdi_data_clean$density,0.002) & cdi_data_clean$density <= quantile(cdi_data_clean$density,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$crime_rate_1000 >= quantile(cdi_data_clean$crime_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$crime_rate_1000,0.998),]
```


### Model construction

Data used for building model:

```{r}
cdi_model = cdi_data_clean %>% 
  dplyr::select(-id,-cty_state, -cty,-state) %>% 
  mutate(region = factor(region))
```


## Introduction

We analyzed data from the “County Demographic Information” (CDI) data set, which contains characteristics of 440 counties in the United States collected from 1990-1992. The primary goal of this investigation is to develop insight relevant to predicting the crime rate in counties. 

## Crime Rate Correlation

```{r}
cdi_cor = cdi_data %>% 
  select(-id,-cty_state, -cty, -state) %>% 
  cor() 
 
corrplot(cdi_cor)
```

According to the plot above, we can see the poverty (Percent below poverty level), beds(Number of hospital beds), docs (Number of active physicians) have positive relationship with crime rate.

Percent high school graduates (Percent of persons 25 years old or older twho completed 12 or more years of school) show a light negative relationship with crime rate.


## Variable Information

```{r, echo = FALSE}
var <- c("id", "cty", "state", "area",	"pop", "pop18",	"pop65",	"docs",	"beds",	"crimes",	"hsgrad",	"bagrad",	"poverty",	"unemp",	"pcincome",	"totalinc",	"region")

var_meaning <- c("ID number", "
County name", "State name", "Land area", "Total population", "Percent of population aged 18-34", "Percent of population aged 65+", "Number of active physicians", "Number of hospital beds", "Total serious crimes", "Percent high school graduates", "Percent bachelor’s degrees", "Percent below poverty level", "Percent unemployment", "Per capita income", "Total personal income", "Geographic region")

var_info <- data.frame(var, var_meaning)

knitr::kable(var_info)
```

## Crime Rate v.s. Poverty

Poverty: Percent of 1990 total population with income below poverty level

```{r}
ggplot(cdi_data, aes(x = poverty, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_poverty = lm(cdi_data$crime_rate_1000 ~ cdi_data$poverty) 
summary(reg_poverty)
fitted_value =  reg_poverty$fitted.values
reg_poverty %>% broom::tidy()
```

## Crime Rate v.s. Beds

Beds: Total number of beds, cribs, and bassinets during 1990

```{r}
ggplot(cdi_data, aes(x = beds_rate_1000, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_beds = lm(cdi_data$crime_rate_1000 ~ cdi_data$beds_rate_1000) 
summary(reg_beds)
reg_beds %>% broom::tidy()
```

## Crime Rate v.s. Docs

docs: Number of active physicians

```{r}
ggplot(cdi_data, aes(x = docs_rate_1000, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_docs = lm(cdi_data$crime_rate_1000 ~ cdi_data$docs_rate_1000) 
summary(reg_docs)
reg_docs %>% broom::tidy()
```

## Crime Rate v.s. hsgrad

hsgrad: Percent of persons 25 years old or older who completed 12 or more years of school

```{r}
ggplot(cdi_data, aes(x = hsgrad, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_docs = lm(cdi_data$crime_rate_1000 ~ cdi_data$hsgrad) 
summary(reg_docs)
reg_docs %>% broom::tidy()
```

## Crime Rate v.s. pop

pop: Estimated 1990 population

```{r}
ggplot(cdi_data, aes(x = pop, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_pop = lm(cdi_data$crime_rate_1000 ~ cdi_data$pop) 
summary(reg_pop)
reg_pop %>% broom::tidy()
```

## Crime Rate v.s. pop18

pop18: Percent of total population in age range from 18-34

```{r}
ggplot(cdi_data, aes(x = pop18, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_pop18 = lm(cdi_data$crime_rate_1000 ~ cdi_data$pop18) 
summary(reg_pop18)
reg_pop18 %>% broom::tidy()
```

## Crime Rate v.s. pop65

```{r}
ggplot(cdi_data, aes(x = pop65, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_pop65 = lm(cdi_data$crime_rate_1000 ~ cdi_data$pop65) 
summary(reg_pop65)
reg_pop65 %>% broom::tidy()
```

## Crime Rate v.s. Area

Land area measured in square miles

```{r}
ggplot(cdi_data, aes(x = area, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_area = lm(cdi_data$crime_rate_1000 ~ cdi_data$area) 
summary(reg_area)
reg_pop65 %>% broom::tidy()
```

## Crime Rate v.s. bagrad

Percent of persons 25 years old or older with bachelor’s degrees

```{r}
ggplot(cdi_data, aes(x = bagrad, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_bagrad = lm(cdi_data$crime_rate_1000 ~ cdi_data$bagrad) 
summary(reg_bagrad)
reg_bagrad %>% broom::tidy()
```

## Multiple Linear Regression

### Interaction

Does the relationship between the crime_rate_1000 and pcincome vary by region status

poverty  

pcincome


```{r}
# poverty v.s. crime_rate_1000
ggplot(cdi_data) + geom_point(aes(x = poverty, y = hsgrad, alpha = .5)) +
  #geom_smooth(method = "lm", se = FALSE) +
  theme_bw()

#slr
slr = lm(poverty ~ hsgrad, data = cdi_data)
summary(slr)
```

Interaction : the relationship between 2 variables is different (or varies) based on the level of third variable?

Does the relationship between the poverty and crime rate vary by region status

```{r}

ggplot(cdi_model, aes(x = poverty, y = crime_rate_1000, color = region)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact1 = lm(poverty ~ hsgrad*region, data = cdi_data)
summary(interact1)
```

Does the relationship between the unemp and crime rate vary by region status

```{r}

ggplot(cdi_model, aes(x = unemp, y = crime_rate_1000, color = region)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact2 = lm(poverty ~ unemp*region, data = cdi_data)
summary(interact2)
```
interaction of pcincome and region

no obvious interaction between pcincome and region

```{r}
ggplot(cdi_model, aes(x = pcincome, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact3 = lm(crime_rate_1000 ~ pcincome*region, data = cdi_data)
summary(interact3)
```

interaction of poverty and region

```{r}
ggplot(cdi_model, aes(x = poverty, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact4 = lm(crime_rate_1000 ~ poverty*region, data = cdi_data)
summary(interact4)
```

Interpret:

Percent below poverty level: Percent of 1990 total population with income below poverty level
The development and economic of different regions are different, therefore, different regions (Northeast, North Central, South, West) would have different poverty rate.

interaction of beds_rate_1000 and region

```{r}
ggplot(cdi_model, aes(x = beds_rate_1000, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact5 = lm(crime_rate_1000 ~ beds_rate_1000*region, data = cdi_data)
summary(interact5)
```

interaction of bagrad and region

```{r}
ggplot(cdi_model, aes(x = bagrad, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact6 = lm(crime_rate_1000 ~ bagrad*region, data = cdi_data)
summary(interact6)

```

Interaction of poverty and pcincome

According to Census Bureau, the number of persons below the official government poverty level was 33.6 million in 1990, representing 13.5 percent of the Nation's population. Thus, we can use this criteria to divide `poverty` into two category: higher than national poverty rate and lower than national poverty rate.


```{r}
poverty_status = cdi_model %>% 
  mutate(national_poverty = if_else(poverty > 13.5, "higher", "lower"))

ggplot(poverty_status, aes(x = pcincome, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) +
  labs(
    title = "Crime Rate and Per Capita Income by Poverty Status",
    x = "Per Capita Income",
    y = "Crime Rate "
  )

```
unemployment and poverty

```{r}
ggplot(poverty_compare, aes(x = unemp, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) 

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ unemp*national_poverty, data = poverty_compare)
summary(interact7)
```

bagrad: Percent of persons 25 years old or older with bachelor’s degrees
20.8% national

pcincome and bagrad

```{r}
bagrad_status = cdi_model %>% 
  mutate(national_bagrad = if_else(bagrad > 20.8, "higher", "lower"))

ggplot(bagrad_status, aes(x = pcincome, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) +
  ylim(0,150) +
  labs(
    title = "Crime Rate and Per Capita Income by Percent Bachelor’s Degrees Status",
    x = "Per Capita Income",
    y = "Crime Rate"
  )

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ pcincome*national_bagrad, data = bagrad_compare)
summary(interact7)
```

poverty and bagrad

```{r}
ggplot(bagrad_compare, aes(x = poverty, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) 

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ poverty*national_bagrad, data = bagrad_compare)
summary(interact7)
```

unemployment and bagrad

```{r}
ggplot(bagrad_compare, aes(x = unemp, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) +
  ylim(0,150)

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ unemp*national_bagrad, data = bagrad_compare)
summary(interact7)
```
density and region

```{r}
ggplot(cdi_model, aes(x = density, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region))
```

density and poverty

```{r}
ggplot(poverty_compare, aes(x = density, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty))
```
density and bagrad

```{r}
ggplot(bagrad_compare, aes(x = density, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad))
```

beds_rate_1000 and poverty

```{r}
ggplot(poverty_compare, aes(x = beds_rate_1000, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) 
```

beds_rate_1000 and bagrad

```{r}
ggplot(bagrad_compare, aes(x = beds_rate_1000, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad))
```
```{r}
mean(cdi_model$pop18)
pop18_compare = cdi_model %>% 
  mutate(pop18_cdi = if_else(bagrad > 28.58, "higher", "lower"))

ggplot(pop18_compare, aes(x = beds_rate_1000, y = crime_rate_1000, color = pop18_cdi, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = pop18_cdi, color = pop18_cdi))

```

## Assessing Muticollinearity

```{r}
mult_fit_back_without = lm(crime_rate ~ pop18 + bagrad + poverty + pcincome + region + pcarea + pcbeds, data = cdi_pc_out)
# Calculate the variance inflation factor (VIF)
check_collinearity(mult_fit_back_without)
# Remove the variable whose vif is larger than 10 
mult_fit_back_without_vif = lm(crime_rate ~ pop18 + bagrad + pcincome + region + pcarea + pcbeds, data = cdi_pc_out)
mult_fit_back_without_vif 
# Add the interaction 
mult_fit_back_without_int = lm(crime_rate ~ pop18 + bagrad + pcincome + region + pcarea + pcbeds + poverty + pcincome*poverty + pcincome*bagrad + pop18*bagrad, data = cdi_pc_out)
mult_fit_back_without_int
```

