---
title: "YiqunJin"
author: "Yiqun Jin"
date: "11/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(corrplot)
```

## Load Data

```{r}
cdi_df = read.csv("data/cdi.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    crime_rate_1000 = 1000 * crimes/pop) %>% # Compute number of crimes per 1000 people.) 
  select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)
```

## Crime Rate Correlation

```{r}
cdi_cor = cdi_df %>% 
  select(-id,-cty_state, -cty, -state) %>% 
  cor() 
 
corrplot(cdi_cor)
```

According to the plot above, we can see the poverty (Percent below poverty level), beds(Number of hospital beds), docs (Number of active physicians) have positive relationship with crime rate.

Percent high school graduates (Percent of persons 25 years old or older twho completed 12 or more years of school) show a light negative relationship with crime rate.


## Variable Information

```{r}
var <- c("id", "cty", "state", "area",	"pop", "pop18",	"pop65",	"docs",	"beds",	"crimes",	"hsgrad",	"bagrad",	"poverty",	"unemp",	"pcincome",	"totalinc",	"region")

var_meaning <- c("ID number", "
County name", "State name", "Land area", "Total population", "Percent of population aged 18-34", "Percent of population aged 65+", "Number of active physicians", "Number of hospital beds", "Total serious crimes", "Percent high school graduates", "Percent bachelor’s degrees", "Percent below poverty level", "Percent unemployment", "Per capita income", "Total personal income", "Geographic region")

var_info <- data.frame(var, var_meaning)

knitr::kable(var_info)
```

## Crime Rate v.s. Poverty

Poverty: Percent of 1990 total population with income below poverty level

```{r}
ggplot(cdi_df, aes(x = poverty, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_poverty = lm(cdi_df$crime_rate_1000 ~ cdi_df$poverty) 
summary(reg_poverty)
fitted_value =  reg_poverty$fitted.values
reg_poverty %>% broom::tidy()
```

## Crime Rate v.s. Beds

Beds: Total number of beds, cribs, and bassinets during 1990

```{r}
ggplot(cdi_df, aes(x = beds_rate_1000, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_beds = lm(cdi_df$crime_rate_1000 ~ cdi_df$beds_rate_1000) 
summary(reg_beds)
reg_beds %>% broom::tidy()
```

## Crime Rate v.s. Docs

docs (Number of active physicians)

```{r}
ggplot(cdi_df, aes(x = docs_rate_1000, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_docs = lm(cdi_df$crime_rate_1000 ~ cdi_df$docs_rate_1000) 
summary(reg_docs)
reg_docs %>% broom::tidy()
```

## Crime Rate v.s. hsgrad

```{r}
ggplot(cdi_df, aes(x = hsgrad, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_docs = lm(cdi_df$crime_rate_1000 ~ cdi_df$hsgrad) 
summary(reg_docs)
reg_docs %>% broom::tidy()
```



## Abstract

## Introduction

## Methods

## Results

## Conclusion/Discussion