---
title: "YiqunJin"
author: "Yiqun Jin"
date: "11/22/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(corrplot)
library(performance)
library(dplyr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Load Data

```{r}
cdi_data = read.csv("data/cdi.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    density = as.numeric(pop)/as.numeric(area),
    crime_rate_1000 = 1000 * crimes/pop) %>% # Compute number of crimes per 1000 people. 
  dplyr::select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)

knitr::kable(head(cdi_data))
  
```

### Remove outliers and high leverage point

```{r}
# Remove high leverage points

cdi_data_clean = cdi_data[cdi_data$area >= quantile(cdi_data$area,0.002) & cdi_data$area <= quantile(cdi_data$area,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop >= quantile(cdi_data_clean$pop,0.002) & cdi_data_clean$pop <= quantile(cdi_data_clean$pop,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop18 >= quantile(cdi_data_clean$pop18,0.002) & cdi_data_clean$pop18 <= quantile(cdi_data_clean$pop18,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop65 >= quantile(cdi_data_clean$pop65,0.002) & cdi_data_clean$pop65 <= quantile(cdi_data_clean$pop65,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$hsgrad >= quantile(cdi_data_clean$hsgrad,0.002) & cdi_data_clean$hsgrad <= quantile(cdi_data_clean$hsgrad,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$bagrad >= quantile(cdi_data_clean$bagrad,0.002) & cdi_data_clean$bagrad <= quantile(cdi_data_clean$bagrad,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$poverty >= quantile(cdi_data_clean$poverty,0.002) & cdi_data_clean$poverty <= quantile(cdi_data_clean$poverty,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$unemp >= quantile(cdi_data_clean$unemp,0.002) & cdi_data_clean$unemp <= quantile(cdi_data_clean$unemp,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pcincome >= quantile(cdi_data_clean$pcincome,0.002) & cdi_data_clean$pcincome <= quantile(cdi_data_clean$pcincome,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$totalinc >= quantile(cdi_data_clean$totalinc,0.002) & cdi_data_clean$totalinc <= quantile(cdi_data_clean$totalinc,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$docs_rate_1000 >= quantile(cdi_data_clean$docs_rate_1000,0.002) & cdi_data_clean$docs_rate_1000 <= quantile(cdi_data_clean$docs_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$density >= quantile(cdi_data_clean$density,0.002) & cdi_data_clean$density <= quantile(cdi_data_clean$density,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$crime_rate_1000 >= quantile(cdi_data_clean$crime_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$crime_rate_1000,0.998),]
```


### Model construction

Data used for building model:

```{r}
cdi_model = cdi_data_clean %>% 
  dplyr::select(-id,-cty_state, -cty,-state) %>% 
  mutate(region = factor(region))
```

## Abstract

condenses the results into a one-paragraph summary.

## Introduction

We analyzed data from the “County Demographic Information” (CDI) data set, which contains characteristics of 440 counties in the United States collected from 1990-1992. The primary goal of this investigation is to develop insight relevant to predicting the crime rate in counties. 

## Methods

## Results



## Conclusion / Discussion


## Crime Rate Correlation

```{r}
cdi_cor = cdi_data %>% 
  select(-id,-cty_state, -cty, -state) %>% 
  cor() 
 
corrplot(cdi_cor)
```

According to the plot above, we can see the poverty (Percent below poverty level), beds(Number of hospital beds), docs (Number of active physicians) have positive relationship with crime rate.

Percent high school graduates (Percent of persons 25 years old or older twho completed 12 or more years of school) show a light negative relationship with crime rate.


## Variable Information

```{r, echo = FALSE}
var <- c("id", "cty", "state", "area","region",	"pop", "pop18",	"pop65",	"docs_rate_1000",	"beds_rate_1000",	"crime_rate_1000",	"hsgrad",	"bagrad",	"poverty",	"unemp",	"pcincome",	"totalinc",	"region")

var_meaning <- c("ID number", "County name", "State name", "Land area","Population Density" ,"Esimate 1990 population", "Percent of population aged 18-34", "Percent of population aged 65+", "Number of active physicians per 1000 people", "Number of hospital beds per 1000 people", "Number of serious crimes per 1000 people", "Percent high school graduates", "Percent bachelor’s degrees", "Percent below poverty level", "Percent unemployment", "Per capita income", "Total personal income", "Geographic region")

var_info <- data.frame(var, var_meaning)

knitr::kable(var_info)
```




```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%",
  echo = TRUE, warning = FALSE, message = FALSE,
  dpi=300
  
)
```

```{r library,warning=FALSE,message=FALSE}
library(tidyverse)
library(corrplot)
library(leaps)
library(performance)
library(MASS)
library(caret)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%",
  echo = TRUE, warning = FALSE, dpi=300
  
)



```

## Step 1: Data Preprocessing

After importing the csv file containing the County Demographic Information (CDI) data, we notice that crimes, physicians, and hospital beds are given as numbers, while other info are given as proportions. We therefore compute the number of crimes, physicians, and hospital beds per 1000 people.

```{r preprocess, messgae = FALSE}
cdi_data = read_csv("./data/cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    density = as.numeric(pop)/as.numeric(area),
    crime_rate_1000 = 1000 * crimes/pop) %>% # Compute number of crimes per 1000 people. 
  dplyr::select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)

knitr::kable(head(cdi_data))
```


## Step 2 - Exploratory Analysis

We then take a closer look of each variables, calculate the pairwise correlations between variables, and list all the correlations between the crime rate (our interest) and all other variables.

```{r exploration}
cdi_data_exp = cdi_data %>%
  dplyr::select(-id,-cty,-state, -cty_state) 

cdi_data_original = cdi_data_exp

cdi_data_exp = cdi_data_exp %>%
  mutate(
    area = ifelse(area %in% boxplot.stats(cdi_data_exp$area)$out,NA,area),
    pop = ifelse(pop %in% boxplot.stats(cdi_data_exp$pop)$out,NA,pop),
    crime_rate_1000 = ifelse(crime_rate_1000 %in% boxplot.stats(cdi_data_exp$crime_rate_1000)$out,NA,crime_rate_1000),
    poverty = ifelse(poverty %in% boxplot.stats(cdi_data_exp$poverty)$out,NA,poverty)
  ) %>%
  na.omit()
```


```{r, fig.width=20,fig.height=20,fig.cap="\\label{fig:figs}boxplot of continuous variables distribution"}
par(mfrow=c(4,3))
boxplot(cdi_data_exp$area,main="Area")
boxplot(cdi_data_exp$pop,main="Population")
boxplot(cdi_data_exp$pop18,main="Population 18-34")
boxplot(cdi_data_exp$pop65,main="Population 65+")
boxplot(cdi_data_exp$hsgrad,main="Highschool grads")
boxplot(cdi_data_exp$bagrad,main="Bachelor's grads")

#par(mfrow=c(2,3))
boxplot(cdi_data_exp$poverty,main="Poverty Rate")
boxplot(cdi_data_exp$unemp,main="Unemployment Rate")
boxplot(cdi_data_exp$pcincome,main="Income Per Capita")
boxplot(cdi_data_exp$totalinc,main="Income Total")
boxplot(cdi_data_exp$docs_rate_1000,main="Active Physicians")
boxplot(cdi_data_exp$beds_rate_1000,main="Hospital Beds")
```

```{r,fig.cap="\\label{fig:figs}Histogram of catagorical variable:region distribution"}
par(mfrow=c(1,1))

ggplot(cdi_data,aes(region)) + 
  geom_histogram(binwidth = 0.5) +
  theme_classic() +
  xlab("Region")+
  ylab("Count")
```


```{r, fig.cap="\\label{fig:figs}boxplot of dependent variable: crime rate"}
boxplot(cdi_data_exp$crime_rate_1000,main="Crime Rate",horizontal = TRUE)
```


```{r, result = "hide"}
# data exploratory
pairs(cdi_data_exp)
```


```{r, fig.cap="\\label{fig:figs}Correlation heatmap"}
# correlation plot
cdi_data_cor = cor(cdi_data_exp)
corrplot(cdi_data_cor, type = "upper", diag = FALSE)
```


```{r}
crime_1000_cor = data.frame(cdi_data_cor) %>% 
  dplyr::select("Crime Rate (Per 1000)" = crime_rate_1000) %>% 
  t()

#knitr::kable(crime_1000_cor,digits = 2) 

```


## Remove outliers and high leverage point

```{r}
# Remove high leverage points

cdi_data_clean = cdi_data[cdi_data$area >= quantile(cdi_data$area,0.002) & cdi_data$area <= quantile(cdi_data$area,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop >= quantile(cdi_data_clean$pop,0.002) & cdi_data_clean$pop <= quantile(cdi_data_clean$pop,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop18 >= quantile(cdi_data_clean$pop18,0.002) & cdi_data_clean$pop18 <= quantile(cdi_data_clean$pop18,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pop65 >= quantile(cdi_data_clean$pop65,0.002) & cdi_data_clean$pop65 <= quantile(cdi_data_clean$pop65,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$hsgrad >= quantile(cdi_data_clean$hsgrad,0.002) & cdi_data_clean$hsgrad <= quantile(cdi_data_clean$hsgrad,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$bagrad >= quantile(cdi_data_clean$bagrad,0.002) & cdi_data_clean$bagrad <= quantile(cdi_data_clean$bagrad,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$poverty >= quantile(cdi_data_clean$poverty,0.002) & cdi_data_clean$poverty <= quantile(cdi_data_clean$poverty,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$unemp >= quantile(cdi_data_clean$unemp,0.002) & cdi_data_clean$unemp <= quantile(cdi_data_clean$unemp,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$pcincome >= quantile(cdi_data_clean$pcincome,0.002) & cdi_data_clean$pcincome <= quantile(cdi_data_clean$pcincome,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$totalinc >= quantile(cdi_data_clean$totalinc,0.002) & cdi_data_clean$totalinc <= quantile(cdi_data_clean$totalinc,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$docs_rate_1000 >= quantile(cdi_data_clean$docs_rate_1000,0.002) & cdi_data_clean$docs_rate_1000 <= quantile(cdi_data_clean$docs_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$beds_rate_1000 >= quantile(cdi_data_clean$beds_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$beds_rate_1000,0.998),]
cdi_data_clean = cdi_data_clean[cdi_data_clean$density >= quantile(cdi_data_clean$density,0.002) & cdi_data_clean$density <= quantile(cdi_data_clean$density,0.998),]

cdi_data_clean = cdi_data_clean[cdi_data_clean$crime_rate_1000 >= quantile(cdi_data_clean$crime_rate_1000,0.002) & cdi_data_clean$beds_rate_1000 <= quantile(cdi_data_clean$crime_rate_1000,0.998),]


```


```{r, fig.cap="\\label{fig:figs}Boxplot of each continuous variables aftern cleaning outliers"}

par(mfrow=c(3,4))
boxplot(cdi_data_clean$area,main="Area")
boxplot(cdi_data_clean$pop,main="Population")
boxplot(cdi_data_clean$pop18,main="Population 18-34")
boxplot(cdi_data_clean$pop65,main="Population 65+")
boxplot(cdi_data_clean$hsgrad,main="Highschool grads")
boxplot(cdi_data_clean$bagrad,main="Bachelor's grads")

boxplot(cdi_data_clean$poverty,main="Poverty Rate")
boxplot(cdi_data_clean$unemp,main="Unemployment Rate")
boxplot(cdi_data_clean$pcincome,main="Income Per Capita")
boxplot(cdi_data_clean$totalinc,main="Income Total")
boxplot(cdi_data_clean$docs_rate_1000,main="Active Physicians")
boxplot(cdi_data_clean$beds_rate_1000,main="Hospital Beds")

```


## Training/Test set split

```{r}
cdi_data_clean = cdi_data_clean %>% 
  dplyr::select(-id,-cty_state, -cty,-state) %>% 
  mutate(region = factor(region))

set.seed(1)
dt = sort(sample(nrow(cdi_data_clean), nrow(cdi_data_clean)*.9))
train_data = cdi_data_clean[dt,]
test_data = cdi_data_clean[-dt,]

```

## Model construction

Data used for building model:

```{r}

cdi_model = train_data

```


### Stepwise regression

```{r}

full.fit = lm(crime_rate_1000 ~ ., data = cdi_model)
summary(full.fit) %>% 
  broom::tidy() %>%
  mutate(p_rank = rank(p.value))

backward = step(full.fit, direction='backward') %>%  broom::tidy() %>%  rename(backward = "term")


both = step(full.fit, direction = "both") %>% broom::tidy() %>% rename(stepwise = "term")


```

Variables chosen from stepwise regression:

```{r}
bind_cols(backward[-1,1],both[-1,1]) %>% knitr::kable()

```

### Criteria based selection


```{r}
sb = regsubsets(crime_rate_1000 ~ ., data = cdi_model, nvmax = 14)
sumsb = summary(sb) # pop pop18 hsgrad bagrad poverty pcincome totalinc region beds_rate_1000 density
```

```{r}
coef(sb, id = 12)

```


```{r, fig.cap="\\label{fig:figs}Subset selection for best parameter numbers"}
par(mfrow=c(1,2))
plot(2:15, sumsb$cp, xlab="No. of parameters", ylab="Cp Statistic") 
abline(0,1)

plot(2:15, sumsb$adjr2, xlab="No of parameters", ylab="Adj R2")
```

According to the output, we determine that the number of variables should be above 12 because $C_p \leq p$. Based on this analysis, we find that `unemp` could also be selected.

### Discussion

We need to remove totalinc, because it can be replaced. totalinc = pcincome * pop.

## Model building from the vairables we selected

```{r}
fit_nest = lm(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density, data = cdi_model)
summary(fit_nest)
```

```{r, fig.cap="\\label{fig:figs}Diagnose plots"}
par(mfrow = c(2,2))
plot(fit_nest)
```


```{r, }
boxcox(fit_nest)


```

box close to 0.5~1. Try y^0.5 transformation

### transformation

```{r}

cdi_model_trans = cdi_model %>% 
  mutate(
    y_sqrt = sqrt(crime_rate_1000)
  )

fit_nest_trans = lm(y_sqrt ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density, data = cdi_model_trans)
summary(fit_nest_trans)

# diagnose
par(mfrow = c(2,2))
plot(fit_nest_trans)



```

Transformed is worse.


Our first model: $$crime\_rate\_1000 = pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds\_rate\_1000 + density$$
                  
## Add Interaction term: poverty+income

According to Census Bureau, the number of persons below the official government poverty level was 33.6 million in 1990, representing 13.5 percent of the Nation's population. Thus, we can use this criteria to divide `poverty` into two category: higher than national poverty rate and lower than national poverty rate.

```{r, fig.cap="\\label{fig:figs}Interaction plot of Income Per Capita and Poverty"}
poverty_status = cdi_model %>% 
  mutate(national_poverty = if_else(poverty > 13.5, "higher", "lower"))

ggplot(poverty_status, aes(x = pcincome, y = crime_rate_1000, color = national_poverty)) + 
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) +
  labs(
    title = "Crime Rate and Per Capita Income by Poverty Status",
    x = "Per Capita Income",
    y = "Crime Rate ",
    color = "Comparison with national avergae"
  )
```


```{r}
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1) %>% broom::tidy()
```



```{r}
check_collinearity(fit_int1)
```

We notice that `density`, `bagrad` are not significant

```{r}
# remove density
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1)

check_collinearity(fit_int1)

# remove bagrad
fit_int1 = lm(crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model)
summary(fit_int1)

check_collinearity(fit_int1)
```


### diagnose
```{r}
par(mfrow = c(2,2))
plot(fit_int1)
boxcox(fit_int1)
```

box close to 0.5~1. Try y^0.5 transformation

### transformation

```{r}

cdi_model_trans = cdi_model %>% 
  mutate(
    y_sqrt = sqrt(crime_rate_1000)
  )

fit_int1_trans = lm(y_sqrt ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model_trans)
summary(fit_int1_trans)

# diagnose
par(mfrow = c(2,2))
plot(fit_int1_trans)




```

Transformed is worse.

The second model: crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome

## Add interaction term: pcincome + bagrad

According to Census Bureau, national percent of persons 25 years old or older with bachelor’s degrees is 20.8%. Thus, we can use this criteria to divide `bagrad` into two category: higher than national `bagrad` and lower than national `bargrad`.

```{r, fig.cap="\\label{fig:figs}Interaction plot of Income Per Capita and Bachelor's Degree Status"}
bagrad_status = cdi_model %>% 
  mutate(national_bagrad = if_else(bagrad > 20.8, "higher", "lower"))

ggplot(bagrad_status, aes(x = pcincome, y = crime_rate_1000, color = national_bagrad)) + 
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) +
  ylim(0,150) +
  labs(
    title = "Crime Rate and Per Capita Income by Percent Bachelor’s Degrees Status",
    x = "Per Capita Income",
    y = "Crime Rate",
    color = "Comparison with national avergae"
  )

```

```{r}
fit_int2 = lm(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad, data = cdi_model)
summary(fit_int2)

check_collinearity(fit_int2)
```

### diagnose

```{r}

par(mfrow = c(2,2))
plot(fit_int2)
boxcox(fit_int2)
```


box close to 0.5~1. Try y^0.5 transformation


### transformation
```{r}

fit_int2_trans = lm(y_sqrt ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad, data = cdi_model_trans)
summary(fit_int2_trans)

# diagnose
par(mfrow=c(2,2))
plot(fit_int2_trans, which = 1) # worse
plot(fit_int2_trans, which = 2) 



```

Transformed is worse.

The third model: crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad
     
                  
## Cross validation

### model 1

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train1 = train(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density,data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)
print(model_train1)

```


### model 2

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train2 = train(crime_rate_1000 ~  
                   pop + pop18 + 
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000  +
                  poverty*pcincome, data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

summary(model_train2)


```

### model 3

```{r}
set.seed(1)
train = trainControl(method = "cv", number = 5)

model_train3 = train(crime_rate_1000 ~  
                  pop + pop18 + bagrad +
                  poverty + unemp + pcincome + pcincome*pop + region +
                  beds_rate_1000 + density +
                  pcincome*bagrad,  data = cdi_model,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)
summary(model_train3)

```

### Compare RMSE

```{r}

model <- c("1", "2", "3")

RMSE <- c(round(model_train1$results$RMSE, 2),  round(model_train2$results$RMSE,2),
          round(model_train3$results$RMSE, 2))

R_sq <- c(round(model_train1$results$Rsquared, 3),
          round(model_train2$results$Rsquared, 3),
          round(model_train3$results$Rsquared, 3))

RMSE_table <- data.frame(model, RMSE, R_sq)

coefs_1 = model_train1$finalModel$coefficients
names_1 = model_train1$finalModel$xNames

knitr::kable(RMSE_table)
```

The second model has the lowest RMSE.

### Model Assessment on testing set

```{r}

test_data = test_data %>%
  mutate(
    y = crime_rate_1000,
    y_model_1 = predict(model_train1,test_data),
    y_model_2 = predict(model_train2,test_data),
    y_model_3 = predict(model_train3,test_data))

RMSPE_1 = sqrt(mean((test_data$y-test_data$y_model_1)^2))
RMSPE_2 = sqrt(mean((test_data$y-test_data$y_model_2)^2))
RMSPE_3 = sqrt(mean((test_data$y-test_data$y_model_3)^2))



model_assessment = 
  tibble(
    RMSPE_1 = round(RMSPE_1,2),
    RMSPE_2 = round(RMSPE_2,2),
    RMSPE_3 = round(RMSPE_3,2)) %>% 
  pivot_longer(RMSPE_1:RMSPE_3,
               names_to = "model", 
               names_prefix = "RMSPE_",
               values_to = "RMSPE") %>%
  left_join(RMSE_table,by="model") %>%
  dplyr::select(Model=model,R_square = R_sq,RMSE,RMSPE)

knitr::kable(model_assessment)
```

## Crime Rate v.s. Poverty

Poverty: Percent of 1990 total population with income below poverty level

```{r}
ggplot(cdi_data, aes(x = poverty, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_poverty = lm(cdi_data$crime_rate_1000 ~ cdi_data$poverty) 
summary(reg_poverty)
fitted_value =  reg_poverty$fitted.values
reg_poverty %>% broom::tidy()
```

## Crime Rate v.s. Beds

Beds: Total number of beds, cribs, and bassinets during 1990

```{r}
ggplot(cdi_data, aes(x = beds_rate_1000, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_beds = lm(cdi_data$crime_rate_1000 ~ cdi_data$beds_rate_1000) 
summary(reg_beds)
reg_beds %>% broom::tidy()
```

## Crime Rate v.s. Docs

docs: Number of active physicians

```{r}
ggplot(cdi_data, aes(x = docs_rate_1000, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_docs = lm(cdi_data$crime_rate_1000 ~ cdi_data$docs_rate_1000) 
summary(reg_docs)
reg_docs %>% broom::tidy()
```

## Crime Rate v.s. hsgrad

hsgrad: Percent of persons 25 years old or older who completed 12 or more years of school

```{r}
ggplot(cdi_data, aes(x = hsgrad, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_docs = lm(cdi_data$crime_rate_1000 ~ cdi_data$hsgrad) 
summary(reg_docs)
reg_docs %>% broom::tidy()
```

## Crime Rate v.s. pop

pop: Estimated 1990 population

```{r}
ggplot(cdi_data, aes(x = pop, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_pop = lm(cdi_data$crime_rate_1000 ~ cdi_data$pop) 
summary(reg_pop)
reg_pop %>% broom::tidy()
```

## Crime Rate v.s. pop18

pop18: Percent of total population in age range from 18-34

```{r}
ggplot(cdi_data, aes(x = pop18, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_pop18 = lm(cdi_data$crime_rate_1000 ~ cdi_data$pop18) 
summary(reg_pop18)
reg_pop18 %>% broom::tidy()
```

## Crime Rate v.s. pop65

```{r}
ggplot(cdi_data, aes(x = pop65, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_pop65 = lm(cdi_data$crime_rate_1000 ~ cdi_data$pop65) 
summary(reg_pop65)
reg_pop65 %>% broom::tidy()
```

## Crime Rate v.s. Area

Land area measured in square miles

```{r}
ggplot(cdi_data, aes(x = area, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_area = lm(cdi_data$crime_rate_1000 ~ cdi_data$area) 
summary(reg_area)
reg_pop65 %>% broom::tidy()
```

## Crime Rate v.s. bagrad

Percent of persons 25 years old or older with bachelor’s degrees

```{r}
ggplot(cdi_data, aes(x = bagrad, y = crime_rate_1000)) + geom_point(alpha = .5) + geom_smooth(method = "lm", se = FALSE)

reg_bagrad = lm(cdi_data$crime_rate_1000 ~ cdi_data$bagrad) 
summary(reg_bagrad)
reg_bagrad %>% broom::tidy()
```

## Multiple Linear Regression

### Interaction

Does the relationship between the crime_rate_1000 and pcincome vary by region status

poverty  

pcincome


```{r}
# poverty v.s. crime_rate_1000
ggplot(cdi_data) + geom_point(aes(x = poverty, y = hsgrad, alpha = .5)) +
  #geom_smooth(method = "lm", se = FALSE) +
  theme_bw()

#slr
slr = lm(poverty ~ hsgrad, data = cdi_data)
summary(slr)
```

Interaction : the relationship between 2 variables is different (or varies) based on the level of third variable?

Does the relationship between the poverty and crime rate vary by region status

```{r}

ggplot(cdi_model, aes(x = poverty, y = crime_rate_1000, color = region)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact1 = lm(poverty ~ hsgrad*region, data = cdi_data)
summary(interact1)
```

Does the relationship between the unemp and crime rate vary by region status

```{r}

ggplot(cdi_model, aes(x = unemp, y = crime_rate_1000, color = region)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact2 = lm(poverty ~ unemp*region, data = cdi_data)
summary(interact2)
```
interaction of pcincome and region

no obvious interaction between pcincome and region

```{r}
ggplot(cdi_model, aes(x = pcincome, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact3 = lm(crime_rate_1000 ~ pcincome*region, data = cdi_data)
summary(interact3)
```

interaction of poverty and region

```{r}
ggplot(cdi_model, aes(x = poverty, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact4 = lm(crime_rate_1000 ~ poverty*region, data = cdi_data)
summary(interact4)
```

Interpret:

Percent below poverty level: Percent of 1990 total population with income below poverty level
The development and economic of different regions are different, therefore, different regions (Northeast, North Central, South, West) would have different poverty rate.

interaction of beds_rate_1000 and region

```{r}
ggplot(cdi_model, aes(x = beds_rate_1000, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact5 = lm(crime_rate_1000 ~ beds_rate_1000*region, data = cdi_data)
summary(interact5)
```

interaction of bagrad and region

```{r}
ggplot(cdi_model, aes(x = bagrad, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region)) 

# fit model with interaction
interact6 = lm(crime_rate_1000 ~ bagrad*region, data = cdi_data)
summary(interact6)

```

Interaction of poverty and pcincome

According to Census Bureau, the number of persons below the official government poverty level was 33.6 million in 1990, representing 13.5 percent of the Nation's population. Thus, we can use this criteria to divide `poverty` into two category: higher than national poverty rate and lower than national poverty rate.


```{r}
poverty_status = cdi_model %>% 
  mutate(national_poverty = if_else(poverty > 13.5, "higher", "lower"))

ggplot(poverty_status, aes(x = pcincome, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) +
  labs(
    title = "Crime Rate and Per Capita Income by Poverty Status",
    x = "Per Capita Income",
    y = "Crime Rate "
  )

```
unemployment and poverty

```{r}
ggplot(poverty_compare, aes(x = unemp, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) 

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ unemp*national_poverty, data = poverty_compare)
summary(interact7)
```

bagrad: Percent of persons 25 years old or older with bachelor’s degrees
20.8% national

pcincome and bagrad

```{r}
bagrad_status = cdi_model %>% 
  mutate(national_bagrad = if_else(bagrad > 20.8, "higher", "lower"))

ggplot(bagrad_status, aes(x = pcincome, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) +
  ylim(0,150) +
  labs(
    title = "Crime Rate and Per Capita Income by Percent Bachelor’s Degrees Status",
    x = "Per Capita Income",
    y = "Crime Rate"
  )

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ pcincome*national_bagrad, data = bagrad_compare)
summary(interact7)
```

poverty and bagrad

```{r}
ggplot(bagrad_compare, aes(x = poverty, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) 

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ poverty*national_bagrad, data = bagrad_compare)
summary(interact7)
```

unemployment and bagrad

```{r}
ggplot(bagrad_compare, aes(x = unemp, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad)) +
  ylim(0,150)

# fit model with interaction
interact7 = lm(crime_rate_1000 ~ unemp*national_bagrad, data = bagrad_compare)
summary(interact7)
```
density and region

```{r}
ggplot(cdi_model, aes(x = density, y = crime_rate_1000, color = region, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = region, color = region))
```

density and poverty

```{r}
ggplot(poverty_compare, aes(x = density, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty))
```
density and bagrad

```{r}
ggplot(bagrad_compare, aes(x = density, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad))
```

beds_rate_1000 and poverty

```{r}
ggplot(poverty_compare, aes(x = beds_rate_1000, y = crime_rate_1000, color = national_poverty, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_poverty, color = national_poverty)) 
```

beds_rate_1000 and bagrad

```{r}
ggplot(bagrad_compare, aes(x = beds_rate_1000, y = crime_rate_1000, color = national_bagrad, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = national_bagrad, color = national_bagrad))
```
```{r}
mean(cdi_model$pop18)
pop18_compare = cdi_model %>% 
  mutate(pop18_cdi = if_else(bagrad > 28.58, "higher", "lower"))

ggplot(pop18_compare, aes(x = beds_rate_1000, y = crime_rate_1000, color = pop18_cdi, alpha = .5)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F, aes(group = pop18_cdi, color = pop18_cdi))

```

## Assessing Muticollinearity

```{r}
mult_fit_back_without = lm(crime_rate ~ pop18 + bagrad + poverty + pcincome + region + pcarea + pcbeds, data = cdi_pc_out)
# Calculate the variance inflation factor (VIF)
check_collinearity(mult_fit_back_without)
# Remove the variable whose vif is larger than 10 
mult_fit_back_without_vif = lm(crime_rate ~ pop18 + bagrad + pcincome + region + pcarea + pcbeds, data = cdi_pc_out)
mult_fit_back_without_vif 
# Add the interaction 
mult_fit_back_without_int = lm(crime_rate ~ pop18 + bagrad + pcincome + region + pcarea + pcbeds + poverty + pcincome*poverty + pcincome*bagrad + pop18*bagrad, data = cdi_pc_out)
mult_fit_back_without_int
```

