---
title: "Yushan Wang"
output: github_document
---

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
library(leaps)
library(ALSM)
library(performance)
library(caret)
library(MASS)

```

## Variable information:

```{r}
var <- c("id", "cty", "state", "area",	"pop", "pop18",	"pop65",	"docs",	"beds",	"crimes",	"hsgrad",	"bagrad",	"poverty",	"unemp",	"pcincome",	"totalinc",	"region")

var_meaning <- c("ID number", "
County name", "State name", "Land area", "Total population", "Percent of population aged 18-34", "Percent of population aged 65+", "Number of active physicians", "Number of hospital beds", "Total serious crimes", "Percent high school graduates", "Percent bachelor’s degrees
", "Percent below poverty level", "Percent unemployment", "Per capita income", "Total personal income", "Geographic region")

var_info <- data.frame(var, var_meaning)

knitr::kable(var_info)
```

## load data

```{r}
cdi_data = read_csv("./data/cdi.csv") %>%
  janitor::clean_names() %>%
  mutate(
    cty_state = str_c(cty,",",state),
    docs_rate_1000 = 1000 * docs/pop, # Compute number of doctors/hospital beds per 1000 people.
    beds_rate_1000 = 1000 * beds/pop,
    density = as.numeric(pop)/as.numeric(area),
    crime_rate_1000 = 1000 * crimes/pop) %>% # Compute number of crimes per 1000 people. 
  dplyr::select(-docs,-beds,-crimes) %>%
  relocate(id,cty_state,cty)


  
```


```{r}
cdi_model = cdi_data %>% 
  dplyr::select(-id,-cty_state, -cty,-state) %>% 
  mutate(region = factor(region))
```


## Criteria based selection


```{r}

sb = regsubsets(crime_rate_1000 ~ ., data = cdi_model, nvmax = 14)
sumsb = summary(sb) # pop pop18 hsgrad bagrad poverty pcincome totalinc region beds_rate_1000 density
sumsb

# plot of Cp and Adj-R2 as functions of parameters
par(mfrow=c(1,2))
plot(2:15, sumsb$cp, xlab="No. of parameters", ylab="Cp Statistic") 
abline(0,1)

plot(2:15, sumsb$adjr2, xlab="No of parameters", ylab="Adj R2")


```

remove totalinc?

## Model building with interaction
region*poverty
0.2 cut-off

```{r}
fit = lm(crime_rate_1000 ~  
                  pop + pop18 + hsgrad + bagrad + 
                  poverty + pcincome + totalinc + region +
                  beds_rate_1000 + density + 
                  region*poverty , data = cdi_model)
summary(fit)
anova(fit)
```

### collinearity
```{r}
check_collinearity(fit)
```

## Model building without interaction

```{r}
fit_nest = lm(crime_rate_1000 ~  
                  pop + pop18 + 
                  poverty + pcincome + totalinc + region +
                  beds_rate_1000 + density, data = cdi_model)
summary(fit_nest)
anova(fit_nest)
```

remove hsgrad, then ba?

## anova

```{r}
# compare nested (small vs large) models
# Ho: smaller model is defensible
anova(fit_nest, fit)
# choose nest model(without region*poverty interaction)
```

## test

```{r}
fit_test = lm(crime_rate_1000 ~  
                  pop + pop18   
                  + pcincome + totalinc + region +
                  beds_rate_1000 + density + 
                  poverty*pcincome, data = cdi_model)
summary(fit_test)
anova(fit_test)

check_collinearity(fit_test)

# diagnose

plot(fit_test, which = 1)
plot(fit_test, which = 2)
boxcox(fit_test)

# transformation

cdi_model_trans = cdi_model %>% 
  mutate(
    y_sqrt = sqrt(crime_rate_1000)
  )

fit_trans = lm(y_sqrt ~  
                  pop + pop18   
                  + pcincome + totalinc + region +
                  beds_rate_1000 + density + 
                  poverty*pcincome, data = cdi_model_trans)

# diagnose
plot(fit_trans, which = 1)
plot(fit_trans, which = 2) # better
boxcox(fit_trans)

# choose transformed!


# model 1!
```

pcincome*bagrad significant
poverty*income significnat, hsgrad, bagrad, poverty insignificant

remove outlier
```{r}
plot(fit_test, which = 4) 

cdiOut = cdi_model[-c(1,6),]

# fit model with and without influential points
with = fit_test 

without = lm(crime_rate_1000 ~  
                  pop + pop18   
                  + pcincome + totalinc + region +
                  beds_rate_1000 + density + 
                  poverty*pcincome, data = cdiOut)

summary(with); summary(without) 

par(mfrow = c(2,2))
plot(without, which = 1)
plot(without, which = 2)

boxcox(without)

# use without
# transformation

cdiOut_trans = cdiOut %>% 
  mutate(
    y_sqrt = sqrt(crime_rate_1000)
  )

fit_out_trans = lm(y_sqrt ~  
                  pop + pop18   
                  + pcincome + totalinc + region +
                  beds_rate_1000 + density + 
                  poverty*pcincome, data = cdiOut_trans)

# diagnose
plot(fit_out_trans, which = 1)
plot(fit_out_trans, which = 2) 
boxcox(fit_out_trans)

# choose transformed!

```

## use new dataset do again


## Try validation

```{r}
set.seed(1)
# Use 5-fold validation and create the training sets
train = trainControl(method = "cv", number = 5)

# Fit the 4-variables model that we discussed in previous lectures
model_train = train(crime_rate_1000 ~  
                  pop + pop18   
                  + pcincome + totalinc + region +
                  beds_rate_1000 + density + 
                  poverty*pcincome, data = cdiOut,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

model_train$finalModel
print(model_train)

#rmse very large :(
```








